---
title: "Scooby Doo Villain Motive Analysis"
author: "Leopoldine Mirtil"
output: rmarkdown::github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Source Data

The data used in this project was uploaded by the user Plummye on [Kaggle.com](https://www.kaggle.com/datasets/williamschooleman/scoobydoo-complete?select=Scooby-Doo+Complete+-+Episode+List+-+Update+9+15+21.csv#). It contains information on all Scooby-Doo episodes, movies and specials as of 9/15/2021. The data has 75 columns covering numerous variables including show/movie information, monster information, culprit information, motives, character appearances,  and gags(jinkies, zoinks, lost glasses, traps set, etc). 

## Objective

I was watching 'Scooby-Doo: Where Are You?'  just as the culprit was being unmasked. This caused me to muse on the motives of most Scooby-Doo villains, especially considering how far they go to achieve their goal. In this analysis, I will examine the motives of the villains from several Scooby-Doo tv shows, featuring the classic gang of Fred, Daphne, Velma, Shaggy and Scooby-Doo.

### Task
  1. Determine the top and bottom five overall motives from all series
  2. Examine the motives between male and female culprits
  3. Identify the top three motives for each show
  4. Identify the top three motives per environment and year

### Step 1: Get Data

#### Load Library

```{r, message=FALSE, warning=FALSE}
#data manipulation
library(tidyr)
library(tidyverse)
library(dplyr)
library(lubridate)

#documentation & markdown
library(knitr)

#creating charts & interactive plots 
library(ggplot2)
library(plotly)
library(webshot) #screenshot of plotly chart in markdown
```


#### Import Data

```{r}
df <- read.csv('Scooby-Doo Completed.csv')
```


### Step 2: Clean & Modify Data

#### Select and Inspect Data

```{r}
#copy data 
sc_df <- df

#select columns
sc_df <- sc_df[,c(2,8,10,42,48,50)]

#view column names 
colnames(sc_df)
```

#### Check for Missing/NA values

```{r}
sapply(sc_df, function(x) sum(is.na(x)))
```


#### Check for Empty/Blank Values

```{r}
sapply(sc_df, function(x) sum(x==""))
```


#### Filter and Modify Data

```{r}
#copy data just in case
scoob <- sc_df

#filter data to: include only the tv shows
                #include only shows with the full Scooby-Doo 
                #remove NULL values from motive column
scoob <- scoob %>%
  filter((format == "TV Series") | (format == "TV Series (segmented)")) %>%
  filter((series.name != "Lego") & (series.name  != "Warner Home Video") &
         (series.name != "Shaggy & Scooby-Doo Get a Clue!") & (series.name  != "Night of the Living Doo") &
         (series.name != "The Scooby-Doo Project") & (series.name  != "The 13 Ghosts of Scooby-Doo") &
         (series.name != "The New Scooby and Scrappy Doo Show") & (series.name  != "Scooby-Doo and Scrappy-Doo (second series)") &
         (series.name != "Scooby-Doo and Scrappy-Doo (first series)") & (series.name  != "Laff-a-Lympics")) %>%
  filter(motive != "NULL")

#change data type 
scoob$date.aired <- as.Date(scoob$date.aired)

#drop format column
scoob <- scoob[-c(3)]

#split column values
scoob <- scoob %>%
  separate_rows(culprit.gender, sep = ",") %>%
  mutate(culprit.gender=trimws(culprit.gender))

#rename columns
scoob <- rename(scoob, series=series.name, environment=setting.terrain)
```


```{r, echo=FALSE}
#export modified file
write.csv(scoob, 'scoob.csv', row.names = FALSE)
```


#### View Data

```{r}
str(scoob)
```


```{r, echo=FALSE}
##clean environment
rm(df, sc_df, scd)
```

### Step 3: Analyze Data

```{r}
summary(scoob)
```

#### Top & Bottom 5 Motives from All Episodes

```{r}
scoob %>%
  group_by(motive) %>%
  summarise(count=n()) %>%
  arrange(desc(count)) %>%
  slice(unique(c(1:5, n() - 4:0))) 
```


#### Top 3 Motives by Series

```{r}
scoob %>%
  group_by(series, motive) %>%
  reframe(count=n()) %>%
  arrange(series, desc(count)) %>%
  group_by(series) %>%
  slice(1:3)
```


#### Motive Count by Gender

```{r}
scoob %>%
  group_by(culprit.gender, motive) %>%
  reframe(count=n()) %>%
  arrange(motive, desc(count)) %>%
  filter(culprit.gender != "NULL")
```


#### Top 3 Motives by Environment

```{r}
scoob %>%
  group_by(environment, motive) %>%
  reframe(count=n()) %>%
  arrange(environment, desc(count)) %>%
  group_by(environment) %>%
  slice(1:3)
```


#### Top 3 Motives by Year

```{r}
scoob %>%
  mutate(aired.year=year(date.aired)) %>%
  group_by(aired.year, motive) %>%
  reframe(count=n()) %>%
  arrange(aired.year, desc(count)) %>%
  group_by(aired.year) %>%
  slice(1:3)
```


### Step 4: Visualize Data

```{r, echo=FALSE, warning=FALSE}
scoob %>%
  group_by(motive) %>%
  summarise(count=n()) %>%
    ggplot(aes(x=fct_reorder(motive, +count), y=count))+
    geom_col(fill="chocolate2") +
    ggtitle('Motive Count of Scooby-Doo Culprits') +
    theme(axis.title.y=element_blank(),
          axis.title.x=element_blank(),
          plot.title = element_text(hjust = 0.5)) +
    coord_flip()
```



```{r, echo=FALSE, warning=FALSE}
scoob %>%
  group_by(culprit.gender, motive) %>%
  filter(culprit.gender != "NULL") %>%
  reframe(count=n()) %>%
  group_by(motive) %>%
  mutate(perc = prop.table(count)) %>%
    ggplot(aes(x=perc, y=reorder(motive, +count), fill=culprit.gender)) +
    geom_bar(position='stack', stat="identity") +
    geom_text(aes(label = scales::percent(perc,accuracy = 1)), size=3,
              position = position_stack(vjust = 0.5,reverse = FALSE)) +
    scale_fill_manual(values=c("mediumorchid4", 
                             "olivedrab3")) +
    scale_x_continuous(labels = scales::percent_format()) +
    ggtitle("Motive Count Percentage by Culprits' Gender") +
    theme(axis.title.x=element_blank(), 
          axis.title.y=element_blank(),
          plot.title = element_text(hjust = .5)) +
    guides(fill=guide_legend(title="Gender")) 
```



```{r, echo=FALSE, warning=FALSE, message=FALSE}
scoob %>%
  mutate(aired.year=year(date.aired)) %>%
  group_by(aired.year, motive) %>%
  reframe(count=n()) %>%
  arrange(aired.year, desc(count)) %>%
  group_by(aired.year) %>%
    ggplot(aes(x=aired.year, y=count, group=motive, fill=motive)) +
    geom_area() +
    geom_line(aes(ymax= count, colour=motive), position="stack", show.legend = FALSE) +
    ggtitle('Annual Motive Count of Scooby-Doo Episodes') +
    theme(axis.title.x=element_blank(), 
          axis.title.y=element_blank(),
          axis.text.x = element_text(angle = 55, hjust=1),
          legend.position = "none",
          plot.title = element_text(hjust = 0.5)) + 
    facet_wrap(vars(motive))
```



```{r, echo=FALSE, warning=FALSE}
scoob %>%
  group_by(series, motive) %>%
  reframe(count=n()) %>%
  arrange(series, desc(count)) %>%
    plot_ly(x = ~motive, y = ~count, color=~series, type = 'bar',
           colors= c("#3399FF", "#D4661B", "#128A84", "#F2C2C2","#8E6357","#B2BB1B","#A51994","#79AF30","#F1D71F")) %>%
      layout(yaxis = list(title = ""),
             xaxis = list(title = "", tickangle = -45),
             margin=list(t=50),
             annotations=list(text="<br>Total Motive Count by Scooby-Doo Series<br>",xref="paper",x=0.80,
                      yref="paper",y=1,yshift=30,showarrow=FALSE,xanchor='center',
                      font=list(size=17)),
             legend = list(x = 110, y = 0.5),
             updatemenus = list(
          list(
          x= 1.65, 
          y=.90,
        buttons = list(
          list(method = "restyle",
               args = list("visible", list(TRUE,TRUE,TRUE,TRUE,TRUE,TRUE,TRUE,TRUE,TRUE)),
               label = "All Series"),
          
          list(method = "restyle",
               args = list("visible", list(TRUE,FALSE,FALSE,FALSE,FALSE,FALSE,FALSE,FALSE,FALSE)),
               label = "A Pup Named Scooby-Doo"),

          list(method = "restyle",
               args = list("visible", list(FALSE,TRUE,FALSE,FALSE,FALSE,FALSE,FALSE,FALSE,FALSE)),
               label = "Be Cool, Scooby-Doo!"),
          
          list(method = "restyle",
               args = list("visible", list(FALSE,FALSE,TRUE,FALSE,FALSE,FALSE,FALSE,FALSE,FALSE)),
               label = "Scooby Doo, Where Are You!"),
          
          list(method = "restyle",
               args = list("visible", list(FALSE,FALSE,FALSE,TRUE,FALSE,FALSE,FALSE,FALSE,FALSE)),
               label = "Scooby-Doo Mystery Incorporated"),
          
          list(method = "restyle",
               args = list("visible", list(FALSE,FALSE,FALSE,FALSE,TRUE,FALSE,FALSE,FALSE,FALSE)),
               label = "Scooby-Doo and Guess Who?"),
          
          list(method = "restyle",
               args = list("visible", list(FALSE,FALSE,FALSE,FALSE,FALSE,TRUE,FALSE,FALSE,FALSE)),
               label = "The New Scooby-Doo Movies"),
          
          list(method = "restyle",
               args = list("visible", list(FALSE,FALSE,FALSE,FALSE,FALSE,FALSE,TRUE,FALSE,FALSE)),
               label = "The New Scooby-Doo Mysteries"),
          
          list(method = "restyle",
               args = list("visible", list(FALSE,FALSE,FALSE,FALSE,FALSE,FALSE,FALSE,TRUE,FALSE)),
               label = "The Scooby-Doo Show"),
          
          list(method = "restyle",
               args = list("visible", list(FALSE,FALSE,FALSE,FALSE,FALSE,FALSE,FALSE,FALSE,TRUE)),
               label = "What's New Scooby-Doo?")))))
```

